---
output:
  html_document: default
---
Peer Assessment 1
==================

In this assignment, we analyze data about personal movement from activity monitoring devices. This first code chunk will download the data and load it into the workspace
```{r loaddata, echo=TRUE}

if(!file.exists("repdata-data-activity.zip")){
        fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
        download.file(url = fileURL, destfile = "repdata-data-activity.zip", method = "curl")
        downloadDate <- date()
        unzip(zipfile = "repdata-data-activity.zip")
        }
data <- read.csv(file = "activity.csv")
```

Next we will do some quick exploratory data analysis
```{r explore, echo=TRUE}
names(data)
str(data)
summary(data)
plot(data$date, data$steps)
```

What is the mean total number of steps taken per day?
```{r meansteps, echo=TRUE}
totalsteps <- aggregate(steps ~ date, data = data, FUN = sum)
hist(totalsteps$steps, main = "Total steps taken each day", xlab = "day", col = "turquoise")
mean(totalsteps$steps)
median(totalsteps$steps)
```

What is the average daily activity pattern?
```{r avgDailyPattern, echo=TRUE}
avgsteps <- aggregate(steps ~ interval, data = data, FUN = mean)
library(ggplot2)
ggplot(avgsteps) + geom_line(aes(x = interval, y = steps)) + geom_smooth(aes(x=interval,y=steps)) + ggtitle(label = "Average number of steps over all days")
maximum <- which(grepl(max(avgsteps$steps), avgsteps$steps))
avgsteps[maximum,]
```
Maximum number of steps seems to be between 8:35am and 8:40am

Imputing missing values
```{r missing, echo=TRUE}
missing <- sum(is.na(data$steps))
missing
```

We'll use the mean step for all days to replace missing values in a new dataset
```{r replaceNA, echo=TRUE}
data2 <- data
na_ID <- which(is.na(data2$steps))
means <- rep(mean(data2$steps, na.rm = TRUE), times = length(na_ID))
data2[na_ID, "steps"] <- means
```

Make a histogram of the total number of steps taken each day, then caluculate mean, median for the total number of steps taken each day
```{r histnewdata, echo=TRUE}
totalstepsnew <- aggregate(steps ~ date, data = data2, FUN = sum)
hist(totalstepsnew$steps, main = "Total number of steps per day for data with no NA", col = "plum")
mean(totalstepsnew$steps)
median(totalstepsnew$steps)
```
After substituting in the mean for NA values, the mean and the median are exactly the same instead of having a difference of 1.19 as in above

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r week-day, echo=TRUE}
library(dplyr)
data2$date <- as.Date(data2$date)
data2 <- mutate(data2, day = weekdays(date)) 
data2$day <- ifelse(data2$day %in% c("Saturday", "Sunday"),"weekend","weekday")
data2[sample(1:dim(data2)[1], size = 5),]
data2$day <- factor(data2$day)
stepsPerDay <- aggregate(steps ~ interval + day, data = data2, mean)
library(lattice)
xyplot(x = steps ~ interval | day, data = stepsPerDay, type = "l", layout = c(1,2), ylab = "Number of steps", xlab = "Interval")
```

Our graphs reveal that fewer steps are taken during weekdays than over weekend days, posibly due to the fact that our anonymous person has more time to walk on weekends